Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.

### 2023-06-05 inst/manuscript/manuscript_microxanox.Rmd
Neraly Final version

| **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |  
| :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
|   C1    | Current code version                                         | v0.9.0                                                                         |  
|   C2    | Permanent links to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
|         |                                                              | [10.5281/zenodo.6624125](https://zenodo.org/record/6624125)                    |
|   C3    | Code Ocean compute capsule                                   |                                                                                |  
|   C4    | Legal Code License                                           | MIT                                                                            |  
|   C5    | Code versioning system used                                  | git                                                                            |  
|   C6    | Software code languages, tools, and services used            | [R](https://cran.r-project.org/index.html)                                     |  
|   C7    | Compilation requirements, operating environments             | [R (>= 4.1.0)](https://cran.r-project.org/index.html)                          |  
|         |                                                              | magrittr                                                                       |  
|         |                                                              | tibble                                                                         |  
|         |                                                              | ggplot2                                                                        |  
|         |                                                              | patchwork                                                                      |  
|         |                                                              | grDevices                                                                      |  
|         |                                                              | stats                                                                          |  
|         |                                                              | mgcv                                                                           |  
|         |                                                              | deSolve                                                                        |  
|         |                                                              | dplyr                                                                          |  
|         |                                                              | tidyr                                                                          |  
|         |                                                              | stringr                                                                        |  
|         |                                                              | multidplyr                                                                     |  
|   C8    | If available Link to developer documentation/manual          | [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) |  
|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Owen.Petchey@ieu.uzh                                       |

@@

We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017. With this goal in mind, we developed the microxanox package [@Krug2022]. The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results. The resulting reproduction is available as one of the package vignettes: [vignette Partial reproduction of Bush et al](https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html).

@@

Here we evidence the impact of the *microxanox* package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems. The first two use cases are described in detail (including the code for reproducing them) in the *User Guide* and the *Partial Reproduction* vignettes. The third is taken from @Limberger2022 and @Petchey2022.

@@

The study of @Bush2017 includes simulations of the effect of oxygen diffusivity (an environmental driver, in the sense that it affects the ecosystem but is not affected by it) on the ecosystem state (oxic or anoxic). The *microxonax* package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity. As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.

@@

```{r uc1, echo=FALSE, fig.align="center", fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the steady states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB\textunderscore1 = cyanobacteria strain 1; SB\\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\\_dynamics()$  function.', message=FALSE, warning=FALSE, include=TRUE, messages=FALSE, out.width=350}
fig <- here::here("inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching.pdf")
if (!file.exists(fig)){
  library(ggplot2)
  library(microxanox)
  library(dplyr)
  x <- readRDS(here::here("inst/manuscript/data/hlh_with_defaults.RDS"))
  p <- plot_dynamics(x)
  p[[1]] <- p[[1]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[2]] <- p[[2]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[3]] <- p[[3]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p_a <- p[[4]]
  p[[4]]$data <- p[[4]]$data %>% dplyr::filter(species != "a")
  p[[4]]$data$species <- p[[4]]$data$species %>%
    dplyr::recode(SO = "Oxidised sulfur", O = "Oxygen", SR = "Sulfide", P = "Phosphorus")
  p[[4]] <- p[[4]] +
    xlab("time (h)") + 
    ylab(expression(atop(log[10](concentration),(µM)))) + 
    labs(colour = "substrate") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p_a$data <- p_a$data %>% dplyr::filter(species == "a")
  p_a$data$species <- p_a$data$species %>%
    dplyr::recode(a = "Oxygen diffusivity ")
  p_a <- p_a +
    ylab(expression(atop(Oxygen~diffusivity,(hr^{-1})))) + 
    labs(colour = "Environmental driver") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p <- p[[1]] / p[[2]] / p[[3]] / p[[4]] / p_a
  ggsave(
    filename = fig,
    plot = p,
    device = "pdf",
    width = 8,
    height = 8
  )  
}
knitr::include_graphics(fig)
```

@@

Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis. The code for this is provided in the *User Guide*.

@@

>>>>

| **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |  
| :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
|   C1    | Current code version                                         | v0.9.1                                                                         |  
|   C2    | Permanent links to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
|         |                                                              | [10.5281/zenodo.7148667](https://zenodo.org/record/7148667)                    |
|   C3    | Code Ocean compute capsule                                   |                                                                                |  
|   C4    | Legal Code License                                           | MIT                                                                            |  
|   C5    | Code versioning system used                                  | [git](https://git-scm.com)                                                                            |  
|   C6    | Software code languages, tools, and services used            | [R](https://cran.r-project.org/index.html)                                     |  
|   C7    | Compilation requirements, operating environments             | [R (>= 4.1.0)](https://cran.r-project.org/index.html)                          |  
|         |                                                              | magrittr                                                                       |  
|         |                                                              | tibble                                                                         |  
|         |                                                              | ggplot2                                                                        |  
|         |                                                              | patchwork                                                                      |  
|         |                                                              | grDevices                                                                      |  
|         |                                                              | stats                                                                          |  
|         |                                                              | mgcv                                                                           |  
|         |                                                              | deSolve                                                                        |  
|         |                                                              | dplyr                                                                          |  
|         |                                                              | tidyr                                                                          |  
|         |                                                              | stringr                                                                        |  
|         |                                                              | multidplyr                                                                     |  
|   C8    | If available Link to developer documentation/manual          | [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) |  
|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Owen.Petchey@ieu.uzh                                       |

@@

We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017. With this goal in mind, we developed the `microxanox` package [@Krug2022]. The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results. The resulting reproduction is available as one of the package vignettes: [vignette Partial reproduction of Bush et al](https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html).

@@

Here we evidence the impact of the *microxanox* package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems. The first two use cases are described in detail (including the code for reproducing them) in the [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) and the *Partial Reproduction* vignettes. The third is taken from @Limberger2022 and @Petchey2022.

@@

The study of @Bush2017 includes simulations of the effect of oxygen diffusivity (an environmental driver, in the sense that it affects the ecosystem but is not affected by it) on the ecosystem state (oxic or anoxic). The `microxonox` package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity. As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.

@@

```{r uc1, echo=FALSE, fig.align="center", fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the steady states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\\_dynamics()$  function.', message=FALSE, warning=FALSE, include=TRUE, messages=FALSE, out.width=350}
fig <- here::here("inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching.pdf")
if (!file.exists(fig)){
  library(ggplot2)
  library(microxanox)
  library(dplyr)
  x <- readRDS(here::here("inst/manuscript/data/hlh_with_defaults.RDS"))
  p <- plot_dynamics(x)
  p[[1]] <- p[[1]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[2]] <- p[[2]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[3]] <- p[[3]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p_a <- p[[4]]
  p[[4]]$data <- p[[4]]$data %>% dplyr::filter(species != "a")
  p[[4]]$data$species <- p[[4]]$data$species %>%
    dplyr::recode(SO = "Oxidised sulfur", O = "Oxygen", SR = "Sulfide", P = "Phosphorus")
  p[[4]] <- p[[4]] +
    xlab("time (h)") + 
    ylab(expression(atop(log[10](concentration),(µM)))) + 
    labs(colour = "substrate") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p_a$data <- p_a$data %>% dplyr::filter(species == "a")
  p_a$data$species <- p_a$data$species %>%
    dplyr::recode(a = "Oxygen diffusivity ")
  p_a <- p_a +
    ylab(expression(atop(Oxygen~diffusivity,(hr^{-1})))) + 
    labs(colour = "Environmental driver") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p <- p[[1]] / p[[2]] / p[[3]] / p[[4]] / p_a
  ggsave(
    filename = fig,
    plot = p,
    device = "pdf",
    width = 8,
    height = 8
  )  
}
knitr::include_graphics(fig)
```

@@

Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis. The code for this is provided in the [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).

@@

# TODO in RMD file
# TODO in LaTeX file
- [ ] fix citation style
- [ ] fix references to figures

### 2023-06-05 inst/manuscript/manuscript_microxanox.tex
Neraly Final version

\title{Microxanox: an R package for simulating an aquatic \(MICR\)obial
ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
\author{true \and true}
\date{2023-06-02}

@@

\pagebreak

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
C1 & Current code version & v0.9.0 \\
C2 & Permanent links to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
& & \href{https://zenodo.org/record/6624125}{10.5281/zenodo.6624125} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & MIT \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used &
\href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments &
\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual &
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide} \\
C9 & Support email for questions &
\href{mailto:Rainer.Krug@uzh.ch}{\nolinkurl{Rainer.Krug@uzh.ch}};
\href{mailto:Owen.Petchey@ieu.uzh}{\nolinkurl{Owen.Petchey@ieu.uzh}} \\
\end{longtable}

@@

We decided to fill this research gap by making a simulation study of how
within functional group biodiversity affects ecosystem responses to
environmental change Limberger et al.
(\protect\hyperlink{ref-Limberger2022}{2022}), and to base our work on
the work and model of Bush et al.
(\protect\hyperlink{ref-Bush2017}{2017}). With this goal in mind, we
developed the microxanox package (\protect\hyperlink{ref-Krug2022}{Krug
and Petchey, 2022}). The first stage of development was to write code
from scratch (as there was no available code to start from) and to
confirm that this new implementation would reproduce the previously
published results. The resulting reproduction is available as one of the
package vignettes:
\href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
Partial reproduction of Bush et al}.

@@

The framework used when writing this package aimed to maximise
simplicity for the user and to make it straightforward to reproduce
results (see the supplement (\protect\hyperlink{ref-Petchey2022}{Petchey
et al., 2022}) to Limberger et al.
(\protect\hyperlink{ref-Limberger2022}{2022}) for an example of how this
is used). As such, all the parameters needed to run a simulation or find
a steady state (i.e.~the final state of the ecosystem) are contained in
a single object (which can easily be created using included functions).
This parameter object is given to a function that runs the simulations
and returns the results. The returned results object is identical to the
parameter object but with an additional slot named \texttt{results},
which contains the simulation results. Thus the returned results object
contains the simulation conditions (parameters) as well as the results,
and can be used to run the simulation again. This promotes
reproducibility and makes incremental changes of individual parameters
with a consecutive re-running of the simulations straightforward.

@@

\caption{Typical flow of a simulation. Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}\label{fig:runsim_example}
\end{figure}

@@

Here we evidence the impact of the \emph{microxanox} package by
describing three use cases and then by describing how the package can be
a starting point for models of other ecosystems. The first two use cases
are described in detail (including the code for reproducing them) in the
\emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
third is taken from Limberger et al.
(\protect\hyperlink{ref-Limberger2022}{2022}) and Petchey et al.
(\protect\hyperlink{ref-Petchey2022}{2022}).

@@

The study of Bush et al. (\protect\hyperlink{ref-Bush2017}{2017})
includes simulations of the effect of oxygen diffusivity (an
environmental driver, in the sense that it affects the ecosystem but is
not affected by it) on the ecosystem state (oxic or anoxic). The
\emph{microxonax} package contains functionality to make a specific
temporal pattern of change in the oxygen diffusivity. As well as
allowing individual simulations during which oxygen diffusivity varies,
this functionality forms the basis of the temporal method for finding
steady states.

@@

\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the steady states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB   extunderscore1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
\end{figure}

@@

Using the package to calculate the extent of hysteresis involves setting
ecosystem and simulation parameters, including parameters for the
finding of steady states across an environmental gradient, running the
steady state finding function, and analysing the results with the
function that calculates extent of hysteresis. The code for this is
provided in the \emph{User Guide}.

@@

As discussed in the Introduction section, the package was motivated by
the question of how biodiversity influences ecosystem responses to
environmental change. Extensive results concerning this question are
given in a separate publication Limberger et al.
(\protect\hyperlink{ref-Limberger2022}{2022}). Here we describe one of
the results, which is that having biodiversity in a functional group can
allow state changes to occur that otherwise would not have. I.e.
biodiversity can qualitatively change the state of the ecosystem.

@@

The \emph{microxanox} R package allows the simulation, visualisation,
and analysis of a model of a microbial ecosystem while allowing
variation in the amount of diversity containing in each of the
functional groups of organisms present. It has been used for the
research described in another paper that describes one of the first
investigations of the effects of diversity on ecosystem resilience
Limberger et al. (\protect\hyperlink{ref-Limberger2022}{2022}). In that
paper, we show that diversity can have large and important effects of
ecosystem responses, highlighting the need for models such as ours, with
which one can easily manipulate the amount of biodiversity. The
\emph{microxanox} package has also been used to reproduce the results of
the paper that inspired the package development
(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}).

@@

\leavevmode\vadjust pre{\hypertarget{ref-Limberger2022}{}}%
Limberger, R., Daugaard, U., Gupta, A., Krug, R., Lemmen, K., Moorsel,
S. van, Suleiman, M., Zuppinger-Dingley, D., Petchey, O., 2022.
Functional diversity can facilitate the collapse of an undesirable
ecosystem state. Authorea Preprint.
doi:\href{https://doi.org/10.22541/au.165513066.66785786/v1}{10.22541/au.165513066.66785786/v1}

>>>>

\title{Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
\author{true \and true}
\date{2023-06-05}

@@

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\pagebreak

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
C1 & Current code version & v0.9.1 \\
C2 & Permanent links to code/repository used for this code version & \url{https://github.com/UZH-PEG/microxanox} \\
& & \href{https://zenodo.org/record/7148667}{10.5281/zenodo.7148667} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & MIT \\
C5 & Code versioning system used & \href{https://git-scm.com}{git} \\
C6 & Software code languages, tools, and services used & \href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments & \href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual & \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} \\
C9 & Support email for questions & \href{mailto:Rainer.Krug@uzh.ch}{\nolinkurl{Rainer.Krug@uzh.ch}}; \href{mailto:Owen.Petchey@ieu.uzh}{\nolinkurl{Owen.Petchey@ieu.uzh}} \\
\end{longtable}

@@

We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change Limberger et al. (\protect\hyperlink{ref-Limberger2022}{2023}), and to base our work on the work and model of Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}). With this goal in mind, we developed the \texttt{microxanox} package (\protect\hyperlink{ref-Krug2022}{Krug and Petchey, 2022}). The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results. The resulting reproduction is available as one of the package vignettes: \href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette Partial reproduction of Bush et al}.

@@

The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement (\protect\hyperlink{ref-Petchey2022}{Petchey et al., 2022}) to Limberger et al. (\protect\hyperlink{ref-Limberger2022}{2023}) for an example of how this is used). As such, all the parameters needed to run a simulation or find a steady state (i.e.~the final state of the ecosystem) are contained in a single object (which can easily be created using included functions). This parameter object is given to a function that runs the simulations and returns the results. The returned results object is identical to the parameter object but with an additional slot named \texttt{results}, which contains the simulation results. Thus the returned results object contains the simulation conditions (parameters) as well as the results, and can be used to run the simulation again. This promotes reproducibility and makes incremental changes of individual parameters with a consecutive re-running of the simulations straightforward.

@@

\caption{Typical flow of a simulation. Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}(\#fig:runsim_example)
\end{figure}

@@

Here we evidence the impact of the \emph{microxanox} package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems. The first two use cases are described in detail (including the code for reproducing them) in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} and the \emph{Partial Reproduction} vignettes. The third is taken from Limberger et al. (\protect\hyperlink{ref-Limberger2022}{2023}) and Petchey et al. (\protect\hyperlink{ref-Petchey2022}{2022}).

@@

The study of Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}) includes simulations of the effect of oxygen diffusivity (an environmental driver, in the sense that it affects the ecosystem but is not affected by it) on the ecosystem state (oxic or anoxic). The \texttt{microxonox} package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity. As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.

@@

\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the steady states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB   extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
\end{figure}

@@

Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis. The code for this is provided in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.

@@

As discussed in the Introduction section, the package was motivated by the question of how biodiversity influences ecosystem responses to environmental change. Extensive results concerning this question are given in a separate publication Limberger et al. (\protect\hyperlink{ref-Limberger2022}{2023}). Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have. I.e. biodiversity can qualitatively change the state of the ecosystem.

@@

The \emph{microxanox} R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present. It has been used for the research described in another paper that describes one of the first investigations of the effects of diversity on ecosystem resilience Limberger et al. (\protect\hyperlink{ref-Limberger2022}{2023}). In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity. The \emph{microxanox} package has also been used to reproduce the results of the paper that inspired the package development (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}).

@@

\leavevmode\vadjust pre{\hypertarget{ref-Limberger2022}{}}%
Limberger, R., Daugaard, U., Gupta, A., Krug, R.M., Lemmen, K.D., Van Moorsel, S.J., Suleiman, M., Zuppinger-Dingley, D., Petchey, O.L., 2023. Functional diversity can facilitate the collapse of an undesirable ecosystem state. Ecology Letters 26, 883--895. doi:\href{https://doi.org/10.1111/ele.14217}{10.1111/ele.14217}
